%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
System Engineering Approach to Extending Endurance of Cooperative Gliders
}


\author{Nahum Camacho$1^{1}$, Vladimir Dobrokhodov$^{2}$, Kevin Jones$^{3}$, and Isaac Kaminer$^{4}$% <-this % stops a space
\thanks{$^{1}$N. Camacho is a graduate student at the Department  of Mechanical and Aerospace Engineering,
        Naval Postgraduate School, Monterey, CA, 93943
        {\tt\small kaminer@nps.edu}}%
\thanks{$^{2}$V.N. Dobrokhodov and $^{3}$K.D. Jones are Research Associate Professors at the Department  of Mechanical and Aerospace Engineering,
        Naval Postgraduate School, Monterey, CA, 93943
        {\tt\small {vldobr, kdjones@nps.edu}}}%
\thanks{$^{4}$I.I. Kaminer is a Professor at the Department  of Mechanical and Aerospace Engineering,
        Naval Postgraduate School, Monterey, CA, 93943
        {\tt\small kaminer@nps.edu}}%
}


\begin{document}



\maketitle
\thispagestyle{empty}
\pagestyle{empty}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

In this paper, the multidisciplinary total-system design approach utilized for the development of a fleet of cooperative, long-endurance gliders is described. The flock is able to harvest energy from the environment both through photo-voltaic energy generation and through exploitation of natural convective lift in the environment, and acts cooperatively to meet mission requirements and to share knowledge of the local environment. The paper begins with a brief overview of the total-energy approach required for such a feat, along with a short description of key system components. This is followed by details of the evolution of a previously-developed architecture that supported autonomous thermaling, to an architecture that considers the total energy budget in all flight segments, and optimizes flight trajectories to maximize energy capture while simultaneously meeting mission goals.


\end{abstract}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

Who is working in the area? What fundamental results were developed and demonstrated in flight? Why the evolution of the previously built architecture is necessary? What are the key building components?

\subsection{Total-Energy System}

Arguably, all aircraft are systems of systems, where all sub-systems interact and affect each other on some level. However, to achieve success in the current project, the inter-dependence of the sub-systems is extreme, such that, in a sense, the over all system design must be solved simultaneously to have any chance of success. The total energy of the system, and the rate of energy capture or dissipation drive the flight trajectory to a large extent. Payload weight turns out to be far less significant than payload power, although both power and weight are interchangeable through induced drag. Glide performance for all feasible flight speeds must be known, and flight controls must utilize this knowledge to optimize glide efficiency for all flight segments.

The energy of the aircraft is governed by the energy equation, increased by energy coming into the system, and decreased by energy leaving the system. There are two methods whereby energy may come into the system: through photovoltaic energy conversion from the sun, and through exploitation of natural convective lift in the environment. Photovoltaic energy production is limited by the size of the array, efficiency of the cells, incident angle of the array to the sun (which couples flight trajectory to solar-power intake), and local solar radiation. Solar radiation is coupled to time of day, geographic location as well as other factors like air quality and altitude. Through a 24 hour period, there is really a small subset of hours where solar radiation is sufficient for any meaningful energy intake. Frequently the energy available from solar radiation is treated as an integration over a 24-hour period. For example, in Monterey California, in June of an average year, a little over 7 $kWh/m^2$ is expected for an array that is flat (not tilted toward the incident radiation).

The second form of energy intake, natural convective lift, is somewhat more difficult to characterize, but was the focus of several previous papers by the present authors and others. In previous developments, a flight control algorithm was developed and tested that allowed multiple, cooperative gliders to locate and utilize thermals for lift in order to stay aloft for extended periods with no propulsion requirement. This process involves a search routine to locate potential thermals, where the cooperation plays a key role, and a thermal-centering algorithm to make sure the glider stays in the thermal as it climbs and as the thermal convects with local winds. While others have performed autonomous thermaling (cite refs), it is believed that the present authors (with Klas Andersson) were the first to do it cooperatively. The cooperation reduces the time needed to locate a potential thermal, and time is energy, as the aircraft is bleeding energy in the form aerodynamic drag and loss of elevation as it glides.

The loss of energy from the system can come in many forms. As mentioned above, during all flight segments aerodynamic drag removes energy from the system, both through viscous and induced drag. Induced drag couples the weight of the system to energy loss. The aircraft may use an optional motor for propulsion, exchanging stored energy for kinetic and/or potential energy, but at a loss due to motor, gear, propeller, and electrical inefficiencies. Additional energy loss may arise from sink, the opposite of convective lift, where the glider flies through a pocket of air that is moving downward. Detection and avoidance of these regions is key. The other primary energy loss is electrical power required for avionics and payload systems. Consideration of a baseline electrical energy requirement along with un-aided glide performance dictates a minimum energy storage requirement to \emph{survive the night}. However, energy storage can typically take two forms, onboard energy storage in the form of batteries, or potential energy in the form of elevation above ground. Early attempts at perpetual solar-powered flight utilized no batteries at all, with only potential energy for storage (cite Sunrise).

More details about these subsystems are given in the following sections that involve selection of the airframe, solar cells, propulsion and energy storage components.

\section{System Level Architecture}
A diagram of the proposed architecture goes here. Discussion of why the components are necessary should be provided.

\subsection{Electric Energy Management Subsystem}

Kevin describes the architecture here: diagram, hardware components, wiring e.t.c. Choice of solar cells. Choice of batteries. Experimental setups to verify the energy density claims. Discussion of the experimental results. It's impact on the future control strategies to sustain overnight flight

\subsection{Potential Energy Management}

Control stuff goes here

\section{HIL and SIL setup}

Condor and Condor API

\section{Preliminary Flight Test Results}

System identification => sink polar=> ID of a thermal=> Thermaling guidance=> Thermal mapping=> Navigation for the purpose of mission goals.

\section{Future steps}

Our plans for the November-February time frame.

\section{CONCLUSIONS}

A conclusion  might elaborate on the importance of the work or suggest applications and extensions.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

Appendixes should appear before the acknowledgment.

\section*{ACKNOWLEDGMENT}

The preferred spelling of the word “acknowledgment” in America is without an “e” after the “g”. Avoid the stilted expression, “One of us (R. B. G.) thanks . . .”  Instead, try “R. B. G. thanks”. Put sponsor acknowledgments in the unnumbered footnote on the first page.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.



\begin{thebibliography}{99}

\bibitem{c1} G. O. Young, “Synthetic structure of industrial plastics (Book style with paper title and editor),” 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15–64.
\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123–135.
\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
\bibitem{c4} B. Smith, “An approach to graphs of linear forms (Unpublished work style),” unpublished.
\bibitem{c5} E. H. Miller, “A note on reflector arrays (Periodical style—Accepted for publication),” IEEE Trans. Antennas Propagat., to be publised.
\bibitem{c6} J. Wang, “Fundamentals of erbium-doped fiber amplifiers arrays (Periodical style—Submitted for publication),” IEEE J. Quantum Electron., submitted for publication.
\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
\bibitem{c8} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, “Electron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),” IEEE Transl. J. Magn.Jpn., vol. 2, Aug. 1987, pp. 740–741 [Dig. 9th Annu. Conf. Magnetics Japan, 1982, p. 301].
\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
\bibitem{c10} J. U. Duncombe, “Infrared navigation—Part I: An assessment of feasibility (Periodical style),” IEEE Trans. Electron Devices, vol. ED-11, pp. 34–39, Jan. 1959.
\bibitem{c11} S. Chen, B. Mulgrew, and P. M. Grant, “A clustering technique for digital communications channel equalization using radial basis function networks,” IEEE Trans. Neural Networks, vol. 4, pp. 570–578, July 1993.
\bibitem{c12} R. W. Lucky, “Automatic equalization for digital communication,” Bell Syst. Tech. J., vol. 44, no. 4, pp. 547–588, Apr. 1965.
\bibitem{c13} S. P. Bingulac, “On the compatibility of adaptive controllers (Published Conference Proceedings style),” in Proc. 4th Annu. Allerton Conf. Circuits and Systems Theory, New York, 1994, pp. 8–16.
\bibitem{c14} G. R. Faulhaber, “Design of service systems with priority reservation,” in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3–8.
\bibitem{c15} W. D. Doyle, “Magnetization reversal in films with biaxial anisotropy,” in 1987 Proc. INTERMAG Conf., pp. 2.2-1–2.2-6.
\bibitem{c16} G. W. Juette and L. E. Zeffanella, “Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),” presented at the IEEE Summer power Meeting, Dallas, TX, June 22–27, 1990, Paper 90 SM 690-0 PWRS.
\bibitem{c17} J. G. Kreifeldt, “An analysis of surface-detected EMG as an amplitude-modulated noise,” presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
\bibitem{c18} J. Williams, “Narrow-band analyzer (Thesis or Dissertation style),” Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993.
\bibitem{c19} N. Kawasaki, “Parametric study of thermal and chemical nonequilibrium nozzle flow,” M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
\bibitem{c20} J. P. Wilkinson, “Nonlinear resonant circuit devices (Patent style),” U.S. Patent 3 624 12, July 16, 1990.






\end{thebibliography}




\end{document}
